---
title: "R Notebook"
output: html_notebook
---


```{r}
#===============================================================================
# Optimized Pan-Tissue Pipeline with ANOVA F-Score & GLMNET
#===============================================================================
# Install from source with specific options

library(data.table)
library(glmnet)
library(caret)
library(pROC)
library(ggplot2)
library(doParallel)
library(dplyr)
library(reshape2)
library(cowplot)
library(magick)


registerDoParallel(cores = detectCores() - 2)

# Load data
merged_gtex <- readRDS("data/processed/merged_gtex.rds")
```

#------------------------------------------------------------------------------
# Data Preparation
#------------------------------------------------------------------------------
```{r}
set.seed(123)
get_tissue_sequence <- function(gtex_data) {
  tissue_stats <- gtex_data[!is.na(SMRIN), .N, by = tissue_main][order(-N)]
  return(tissue_stats$tissue_main)
}

tissue_sequence <- get_tissue_sequence(merged_gtex)
n_tissues_to_test <- 29
current_tissues <- tissue_sequence[1:n_tissues_to_test]

# Prepare dataset
tissue_data <- merged_gtex[tissue_main %in% current_tissues & !is.na(SMRIN) & !is.na(SMATSSCR)]
tissue_data$tissue_class <- factor(tissue_data$tissue_main, levels = current_tissues)
tissue_data$rin_quartile <- cut(tissue_data$SMRIN, 
                               breaks = quantile(tissue_data$SMRIN, probs = 0:4/4),
                               include.lowest = TRUE, labels = paste0("Q", 1:4))

# Autolysis are categorical values(0,1,2,3)
tissue_data$autolysis_quartile <- as.factor(paste0("A", tissue_data$SMATSSCR))
tissue_data$strat_var <- paste(tissue_data$tissue_class, tissue_data$rin_quartile, sep = "_")

# Feature preparation
feature_cols <- grep("^feature_", names(tissue_data), value = TRUE)

# Train-test split
train_indices <- createDataPartition(tissue_data$strat_var, p = 0.7, list = FALSE)
train_data <- tissue_data[train_indices]
test_data <- tissue_data[-train_indices]
```

#-----------------------------------------------------------------------------------
# Stage 1: ANOVA-based Feature Selection + Tissue Classification using GLMNET 
#-----------------------------------------------------------------------------------
```{r}
# ANOVA F-test feature selection
pb <- txtProgressBar(min = 0, max = length(feature_cols), style = 3)
f_scores <- sapply(seq_along(feature_cols), function(i) {
  f <- feature_cols[i]
  anova_result <- aov(train_data[[f]] ~ train_data$tissue_class)
  setTxtProgressBar(pb, i)
  return(summary(anova_result)[[1]]$`F value`[1])
})
close(pb)

names(f_scores) <- feature_cols
top_100_features <- names(sort(f_scores, decreasing = TRUE, na.last = TRUE)[1:500])

# Train tissue classifier
train_tissue_classifier_optimized <- function(train_data, features) {
  train_subset <- train_data[, c(features, "tissue_class"), with = FALSE]
  X_train <- as.matrix(train_subset[, ..features])
  y_train <- train_subset$tissue_class
  
  # Create stratified folds
  set.seed(123)
  stratified_folds <- createFolds(y_train, k = 5, list = TRUE, returnTrain = FALSE)
  foldid <- rep(NA, length(y_train))
  for(i in 1:length(stratified_folds)) {
    foldid[stratified_folds[[i]]] <- i
  }
  
  # Cross-validation with timing
  start_time <- Sys.time()
  model <- cv.glmnet(X_train, y_train, 
                     family = "multinomial", 
                     alpha = 1,
                     foldid = foldid,
                     standardize = TRUE,
                     parallel = TRUE)
  
  print(paste("Training completed in", round(as.numeric(difftime(Sys.time(), start_time, units = "mins")), 1), "minutes"))
  
  return(list(
    model = model,
    features = features,
    tissue_levels = levels(y_train)
  ))
}

tissue_classifier <- train_tissue_classifier_optimized(train_data, top_100_features)
```

#------------------------------------------------------------------------------
# Stage 2: Tissue-Specific RIN Models (Lasso)
#------------------------------------------------------------------------------
```{r}
train_rin_models_lasso <- function(train_data, features, tissues) {
  rin_models <- list()
  performance_results <- data.frame()  # Store performance for sorting
  
  for(tissue in tissues) {
    tissue_subset <- train_data[tissue_class == tissue]
    if(nrow(tissue_subset) < 15) {
      performance_results <- rbind(performance_results, 
                                 data.frame(tissue = tissue, n = nrow(tissue_subset), 
                                          r_val = NA, rmse = NA))
      next
    }
    
    # Use ALL features - let Lasso do the selection
    selected_features <- features
    X_tissue <- as.matrix(tissue_subset[, ..selected_features])
    y_tissue <- tissue_subset$SMRIN
    
    # Stratified CV by RIN quartiles
    set.seed(123)
    rin_quartiles <- cut(y_tissue, breaks = quantile(y_tissue, probs = 0:4/4), include.lowest = TRUE)
    folds <- createFolds(rin_quartiles, k = 5, returnTrain = FALSE)
    foldid <- rep(NA, length(y_tissue))
    for(i in seq_along(folds)) foldid[folds[[i]]] <- i
    
    # Built-in CV for RIN models - standardize=TRUE by default
    set.seed(123)
    cv_fit <- cv.glmnet(X_tissue, y_tissue, alpha = 1, foldid = foldid, parallel = TRUE)
    model <- glmnet(X_tissue, y_tissue, alpha = 1, lambda = cv_fit$lambda.min)
    
    # Calculate performance on training data
    pred <- predict(model, X_tissue, s = cv_fit$lambda.min)
    r_val <- cor(pred, y_tissue)
    rmse <- sqrt(mean((pred - y_tissue)^2))
    
    rin_models[[tissue]] <- list(
      model = model,
      lambda = cv_fit$lambda.min,
      features = selected_features
    )
    
    performance_results <- rbind(performance_results, 
                               data.frame(tissue = tissue, n = nrow(tissue_subset), 
                                        r_val = r_val, rmse = rmse))
  }
  
  performance_results <- performance_results[order(-performance_results$r_val, na.last = TRUE), ]
  
  cat("Tissue-Specific RIN Performance:\n")
  cat("================================\n")
  
  for(i in 1:nrow(performance_results)) {
    row <- performance_results[i, ]
    if(is.na(row$r_val)) {
      cat(sprintf("%-20s: Insufficient samples (n=%d)\n", row$tissue, row$n))
    } else {
      cat(sprintf("%-20s: n=%3d, R=%.3f, RMSE=%.3f\n", row$tissue, row$n, row$r_val, row$rmse))
    }
  }
  
  return(list(models = rin_models, performance = performance_results))
}

rin_full_results <- train_rin_models_lasso(train_data, feature_cols, current_tissues)
rin_models <- rin_full_results$models
rin_training_performance <- rin_full_results$performance
```

#------------------------------------------------------------------------------
# Stage 2.1: Tissue-Specific Autolysis Models (Lasso)
#------------------------------------------------------------------------------
```{r}
train_autolysis_models_lasso <- function(train_data, features, tissues) {
  autolysis_models <- list()
  performance_results <- data.frame()
  
  for(tissue in tissues) {
    tissue_subset <- train_data[tissue_class == tissue]
    if(nrow(tissue_subset) < 15) {
      performance_results <- rbind(performance_results, 
                                 data.frame(tissue = tissue, n = nrow(tissue_subset), 
                                          accuracy = NA, r_val = NA, rmse = NA))
      next
    }
    
    # Use ALL features - let Lasso do the selection
    selected_features <- features
    X_tissue <- as.matrix(tissue_subset[, ..selected_features])
    y_tissue <- tissue_subset$SMATSSCR
    
# Stratified CV by autolysis categories
    set.seed(123)
    autolysis_cat <- factor(y_tissue, levels = 0:3)
    if(length(unique(autolysis_cat)) > 1) {
      folds <- createFolds(autolysis_cat, k = 5, returnTrain = FALSE)
      foldid <- rep(NA, length(y_tissue))
      for(i in seq_along(folds)) foldid[folds[[i]]] <- i
      
      set.seed(123)
      cv_fit <- cv.glmnet(X_tissue, y_tissue, alpha = 1, foldid = foldid, parallel = TRUE)
    } else {
      set.seed(123)
      cv_fit <- cv.glmnet(X_tissue, y_tissue, alpha = 1, nfolds = 5, parallel = TRUE)
    }
    
    model <- glmnet(X_tissue, y_tissue, alpha = 1, lambda = cv_fit$lambda.min)
    
    # Calculate performance on training data with zero variance check
    pred <- predict(model, X_tissue, s = cv_fit$lambda.min)
    rmse <- sqrt(mean((pred - y_tissue)^2))
    
    # Handle zero variance cases
    # Handle zero variance cases
    if(var(pred) == 0 || var(y_tissue) == 0) {
      r_val <- NA
    } else {
      r_val <- cor(pred, y_tissue)
    }
    
    autolysis_models[[tissue]] <- list(
      model = model,
      lambda = cv_fit$lambda.min,
      features = selected_features
    )
  # Store performance results
    performance_results <- rbind(performance_results, 
                               data.frame(tissue = tissue, n = nrow(tissue_subset), 
                                        r_val = r_val, rmse = rmse))
  }
  
  # Sort by R value (descending) and print
  performance_results <- performance_results[order(-performance_results$r_val, na.last = TRUE), ]
  
  cat("Tissue-Specific Autolysis Performance:\n")
  cat("======================================\n")
  
  for(i in 1:nrow(performance_results)) {
    row <- performance_results[i, ]
    if(is.na(row$r_val)) {
      cat(sprintf("%-20s: Insufficient samples (n=%d)\n", row$tissue, row$n))
    } else {
      cat(sprintf("%-20s: n=%3d, R=%.3f, RMSE=%.3f\n", row$tissue, row$n, row$r_val, row$rmse))
    }
  }
  
  return(list(models = autolysis_models, performance = performance_results))
}

autolysis_full_results <- train_autolysis_models_lasso(train_data, feature_cols, current_tissues)
autolysis_models <- autolysis_full_results$models
autolysis_training_performance <- autolysis_full_results$performance
```

#------------------------------------------------------------------------------
# Batch Prediction
#------------------------------------------------------------------------------
```{r}
predict_batch_optimized <- function(test_data, tissue_classifier, rin_models, autolysis_models, confidence_threshold = 0.6) {
  # Tissue classification - Fixed for glmnet
  test_matrix <- as.matrix(test_data[, tissue_classifier$features, with = FALSE])
  
  # Get probabilities from glmnet (returns matrix)
  all_probs <- predict(tissue_classifier$model, test_matrix, s = "lambda.min", type = "response")
  if(length(dim(all_probs)) == 3) all_probs <- all_probs[,,1]  # Extract 2D matrix
  
  # Get predicted classes
  predicted_tissues <- apply(all_probs, 1, function(x) tissue_classifier$tissue_levels[which.max(x)])
  confidences <- apply(all_probs, 1, max)
  
  # RIN prediction
  n_samples <- nrow(test_data)
  rin_predictions <- rep(NA, n_samples)
  used_for_rin <- rep(FALSE, n_samples)
  
  for(tissue in names(rin_models)) {
    tissue_mask <- (predicted_tissues == tissue) & (confidences >= confidence_threshold)
    if(any(tissue_mask)) {
      tissue_indices <- which(tissue_mask)
      tissue_samples <- test_data[tissue_indices]
      rin_model <- rin_models[[tissue]]
      
      tissue_matrix <- as.matrix(tissue_samples[, rin_model$features, with = FALSE])
      batch_rin_pred <- predict(rin_model$model, tissue_matrix, s = rin_model$lambda)
      
      rin_predictions[tissue_indices] <- as.numeric(batch_rin_pred)
      used_for_rin[tissue_indices] <- TRUE
    }
  }
  
  # Autolysis prediction - Same logic as RIN
  autolysis_predictions <- rep(NA, n_samples)
  used_for_autolysis <- rep(FALSE, n_samples)
  
  for(tissue in names(autolysis_models)) {
    tissue_mask <- (predicted_tissues == tissue) & (confidences >= confidence_threshold)
    if(any(tissue_mask)) {
      tissue_indices <- which(tissue_mask)
      tissue_samples <- test_data[tissue_indices]
      autolysis_model <- autolysis_models[[tissue]]
      
      tissue_matrix <- as.matrix(tissue_samples[, autolysis_model$features, with = FALSE])
      batch_autolysis_pred <- predict(autolysis_model$model, tissue_matrix, s = autolysis_model$lambda)
      
      autolysis_predictions[tissue_indices] <- as.numeric(batch_autolysis_pred)
      used_for_autolysis[tissue_indices] <- TRUE
    }
  }
  
  # Results
  results <- data.table(
    true_tissue = test_data$tissue_class,
    true_rin = test_data$SMRIN,
    true_autolysis = test_data$SMATSSCR,
    predicted_tissue = factor(predicted_tissues, levels = tissue_classifier$tissue_levels),
    confidence = confidences,
    rin_prediction = rin_predictions,
    autolysis_prediction = autolysis_predictions,
    used_for_rin = used_for_rin,
    used_for_autolysis = used_for_autolysis
  )
  
  # Add probability columns
  for(i in 1:ncol(all_probs)) {
    tissue <- tissue_classifier$tissue_levels[i]
    prob_col <- paste0("prob_", gsub("[^A-Za-z0-9]", "_", tissue))
    results[[prob_col]] <- all_probs[, i]
  }
  
  return(results)
}

test_results <- predict_batch_optimized(test_data, tissue_classifier, rin_models, autolysis_models)
```

#------------------------------------------------------------------------------
# Test Performance Summary
#------------------------------------------------------------------------------
```{r}
# Function to get test performance summary
get_test_performance_summary <- function(test_results) {
  # RIN test performance
  rin_test_perf <- test_results[used_for_rin == TRUE & !is.na(rin_prediction), .(
    test_r = cor(rin_prediction, true_rin),
    test_rmse = sqrt(mean((rin_prediction - true_rin)^2)),
    test_n = .N
  ), by = predicted_tissue][order(-test_r)]
  
    cat("Test Performance - Top 5 RIN:\n")
    print(head(rin_test_perf, 5))
  
  # Autolysis test performance  
  auto_test_perf <- test_results[used_for_autolysis == TRUE & !is.na(autolysis_prediction), .(
    test_r = cor(autolysis_prediction, true_autolysis),
    test_rmse = sqrt(mean((autolysis_prediction - true_autolysis)^2)),
    test_n = .N
  ), by = predicted_tissue][order(-test_r)]
  
  cat("\nTest Performance - Top 5 Autolysis:\n") 
  print(head(auto_test_perf, 5))
  
  return(list(rin = rin_test_perf, autolysis = auto_test_perf))
}

# Get test performance
test_performance <- get_test_performance_summary(test_results)

```

#------------------------------------------------------------------------------
# Performance Evaluation
#------------------------------------------------------------------------------
```{r}
evaluate_performance <- function(test_results, tissues) {
  cm <- confusionMatrix(test_results$predicted_tissue, test_results$true_tissue)
  
  # Calculate AUC for each tissue
  auc_values <- numeric(length(tissues))
  names(auc_values) <- tissues
  
  for(tissue in tissues) {
    binary_labels <- as.numeric(test_results$true_tissue == tissue)
    prob_col <- paste0("prob_", gsub("[^A-Za-z0-9]", "_", tissue))
    if(prob_col %in% names(test_results) && length(unique(binary_labels)) > 1) {
      roc_obj <- roc(binary_labels, test_results[[prob_col]], quiet = TRUE)
      auc_values[tissue] <- as.numeric(auc(roc_obj))
    }
  }
  
  macro_auc <- mean(auc_values, na.rm = TRUE)
  
  # RIN metrics
  rin_subset <- test_results[used_for_rin == TRUE & !is.na(rin_prediction)]
  rin_r <- if(nrow(rin_subset) > 0) cor(rin_subset$rin_prediction, rin_subset$true_rin) else NA
  rin_rmse <- if(nrow(rin_subset) > 0) sqrt(mean((rin_subset$rin_prediction - rin_subset$true_rin)^2)) else NA
  rin_coverage <- nrow(rin_subset) / nrow(test_results)
  
  # Autolysis metrics - Classification approach with all categories
  autolysis_subset <- test_results[used_for_autolysis == TRUE & !is.na(autolysis_prediction)]
  if(nrow(autolysis_subset) > 0) {
    # Use all possible categories (0,1,2,3)
    all_categories <- 0:3
    
    # Round predictions to nearest integer and clamp to 0-3
    autolysis_pred_rounded <- round(autolysis_subset$autolysis_prediction)
    autolysis_pred_rounded <- pmax(0, pmin(3, autolysis_pred_rounded))
    
    # Classification accuracy
    autolysis_accuracy <- mean(autolysis_pred_rounded == autolysis_subset$true_autolysis)
    
    # Correlation with continuous predictions
    autolysis_r <- cor(autolysis_subset$autolysis_prediction, autolysis_subset$true_autolysis)
    autolysis_rmse <- sqrt(mean((autolysis_subset$autolysis_prediction - autolysis_subset$true_autolysis)^2))
    
    # Confusion matrix with all categories (0,1,2,3)
    autolysis_cm <- table(
      Predicted = factor(autolysis_pred_rounded, levels = all_categories), 
      Actual = factor(autolysis_subset$true_autolysis, levels = all_categories)
    )
    
    
  } else {
    autolysis_accuracy <- NA
    autolysis_r <- NA
    autolysis_rmse <- NA
    autolysis_cm <- NULL
  }
  autolysis_coverage <- nrow(autolysis_subset) / nrow(test_results)
  
  return(list(
    accuracy = cm$overall["Accuracy"],
    macro_f1 = mean(cm$byClass[, "F1"], na.rm = TRUE),
    macro_auc = macro_auc,
    auc_values = auc_values,
    rin_r = rin_r,
    rin_rmse = rin_rmse,
    rin_coverage = rin_coverage,
    autolysis_accuracy = autolysis_accuracy,
    autolysis_r = autolysis_r,
    autolysis_rmse = autolysis_rmse,
    autolysis_coverage = autolysis_coverage,
    autolysis_confusion_matrix = autolysis_cm,
    confusion_matrix = cm$table
  ))
}

performance <- evaluate_performance(test_results, current_tissues)
```

#------------------------------------------------------------------------------
# Print overall Testing data performance summary
#------------------------------------------------------------------------------
```{r}
# Print performance summary
cat("=== PERFORMANCE SUMMARY ===\n")
cat("Tissue Classification:\n")
cat(sprintf("  Accuracy: %.3f\n", performance$accuracy))
cat(sprintf("  Macro F1: %.3f\n", performance$macro_f1))
cat(sprintf("  Macro AUC: %.3f\n", performance$macro_auc))
cat("\nRIN Prediction:\n")
cat(sprintf("  Correlation: %.3f\n", performance$rin_r))
cat(sprintf("  RMSE: %.3f\n", performance$rin_rmse))
cat(sprintf("  Coverage: %.3f\n", performance$rin_coverage))
cat("\nAutolysis Prediction:\n")
cat(sprintf("  Accuracy: %.3f\n", performance$autolysis_accuracy))
cat(sprintf("  Correlation: %.3f\n", performance$autolysis_r))
cat(sprintf("  RMSE: %.3f\n", performance$autolysis_rmse))
cat(sprintf("  Coverage: %.3f\n", performance$autolysis_coverage))
```



#------------------------------------------------------------------------------
# Quality Prediction Visualization (RIN + Autolysis)
#------------------------------------------------------------------------------
```{r}
#------------------------------------------------------------------------------
# Simplified Quality Prediction Visualization
#------------------------------------------------------------------------------

# Unified RIN plot function 
create_rin_plot <- function(test_results, top_n = NULL, show_legend = FALSE) {
  rin_subset <- test_results[used_for_rin == TRUE & !is.na(rin_prediction)]
  if(nrow(rin_subset) == 0) return(NULL)
  
  # Filter to top N tissues if specified
  if(!is.null(top_n)) {
    tissue_performance <- rin_subset[, .(
      correlation = cor(rin_prediction, true_rin),
      n_samples = .N
    ), by = predicted_tissue][order(-correlation)]
    
    top_tissues <- head(tissue_performance$predicted_tissue, top_n)
    rin_subset <- rin_subset[predicted_tissue %in% top_tissues]
    
    # Add facet labels with R values
    rin_subset <- merge(rin_subset, tissue_performance[, .(predicted_tissue, correlation)], by = "predicted_tissue")
    rin_subset$facet_label <- paste0(rin_subset$predicted_tissue, "\n(R = ", round(rin_subset$correlation, 3), ")")
    facet_layer <- facet_wrap(~facet_label, scales = "free")
    title_suffix <- paste(" - Top", top_n, "Tissues")
    coord_layer <- NULL  # No coord_fixed for faceted plots
  } else {
    overall_r <- cor(rin_subset$rin_prediction, rin_subset$true_rin)
    facet_layer <- NULL
    title_suffix <- paste0("\nOverall R = ", round(overall_r, 3), " (n = ", nrow(rin_subset), ")")
    coord_layer <- coord_fixed()  # Only use coord_fixed for non-faceted plots
  }
  
  p <- ggplot(rin_subset, aes(x = true_rin, y = rin_prediction)) +
    geom_point(aes(color = predicted_tissue), alpha = 0.6, size = if(is.null(top_n)) 1 else 1.5) +
    geom_abline(intercept = 0, slope = 1, color = "red", linetype = "dashed") +
    geom_smooth(method = "lm", se = TRUE, color = "blue") +
    scale_color_manual(values = tissue_colors, name = "Tissue") +
    facet_layer +
    coord_layer +
    labs(x = "True RIN Score", y = "Predicted RIN Score",
         title = paste0("RIN Prediction Performance", title_suffix)) +
    theme_minimal() +
    theme(legend.position = if(show_legend) "bottom" else "none")
  
  return(p)
}

# Unified Autolysis plot function - FIXED  
create_autolysis_plot <- function(test_results, top_n = NULL, show_legend = FALSE) {
  autolysis_subset <- test_results[used_for_autolysis == TRUE & !is.na(autolysis_prediction)]
  if(nrow(autolysis_subset) == 0) return(NULL)
  
  # Filter to top N tissues if specified
 
  if(!is.null(top_n)) {
    tissue_performance <- autolysis_subset[, .(
      correlation = {
        if(var(autolysis_prediction) == 0 || var(true_autolysis) == 0) {
          NA_real_
        } else {
          cor(autolysis_prediction, true_autolysis)
        }
      },
      n_samples = .N
    ), by = predicted_tissue][order(-correlation, na.last = TRUE)]
    
    top_tissues <- head(tissue_performance$predicted_tissue, top_n)
    autolysis_subset <- autolysis_subset[predicted_tissue %in% top_tissues]
    
    # Add facet labels with R values  
    autolysis_subset <- merge(autolysis_subset, tissue_performance[, .(predicted_tissue, correlation)], by = "predicted_tissue")
    autolysis_subset$facet_label <- paste0(autolysis_subset$predicted_tissue, "\n(R = ", 
                                          ifelse(is.na(autolysis_subset$correlation), "NA", 
                                                round(autolysis_subset$correlation, 3)), ")")
    
    # CRITICAL FIX: Create ordered factor based on correlation rank
    ordered_labels <- tissue_performance[predicted_tissue %in% top_tissues]$predicted_tissue
    ordered_facet_labels <- paste0(ordered_labels, "\n(R = ", 
                                  round(tissue_performance[predicted_tissue %in% top_tissues]$correlation, 3), ")")
    
    autolysis_subset$facet_label <- factor(autolysis_subset$facet_label, 
                                          levels = ordered_facet_labels)
    
    facet_layer <- facet_wrap(~facet_label, scales = "free")
    title_suffix <- paste(" - Top", top_n, "Tissues")
    
    # Use fill colors for top N
    p <- ggplot(autolysis_subset, aes(x = factor(true_autolysis, levels = 0:3), y = autolysis_prediction)) +
      geom_boxplot(outlier.shape = NA, fill = "lightgray", alpha = 0.7) +
      geom_jitter(aes(color = predicted_tissue), width = 0.2, height = 0, alpha = 0.6, size = 1.2) +
      scale_color_manual(values = tissue_colors, name = "Tissue") +
      theme_minimal() +
      theme(legend.position = if(show_legend) "bottom" else "none") 
    
  } else {
    overall_r <- cor(autolysis_subset$autolysis_prediction, autolysis_subset$true_autolysis)
    facet_layer <- NULL
    title_suffix <- paste0("\nOverall R = ", round(overall_r, 3), " (n = ", nrow(autolysis_subset), ")")
    
    # Use tissue colors for all tissues
    p <- ggplot(autolysis_subset, aes(x = factor(true_autolysis, levels = 0:3), y = autolysis_prediction)) +
      geom_boxplot(alpha = 0.7, outlier.shape = NA, fill = "lightblue") +
      geom_jitter(aes(color = predicted_tissue), width = 0.2, alpha = 0.4, size = 0.8) +
      scale_color_manual(values = tissue_colors, name = "Tissue") +
      theme_minimal() +
      theme(legend.position = if(show_legend) "bottom" else "none")
  }
  
  p <- p +
    facet_layer +
    labs(x = "True Autolysis Score", y = "Predicted Autolysis Score",
         title = paste0("Autolysis Prediction Performance", title_suffix)) +
    scale_x_discrete(drop = FALSE)
  
  return(p)
}

# Generate plots using the new unified functions
rin_plot <- create_rin_plot(test_results, top_n = 3)
autolysis_plot <- create_autolysis_plot(test_results, top_n = 3) 
pan_rin_plot <- create_rin_plot(test_results, show_legend = FALSE)
pan_autolysis_plot <- create_autolysis_plot(test_results, show_legend = FALSE)

# Save plots
if(!is.null(rin_plot)) {
  print(rin_plot)
  ggsave("output/figures/rin_top3_scatter.pdf", rin_plot, width = 12, height = 8, dpi = 300)
}

if(!is.null(autolysis_plot)) {
  print(autolysis_plot)
  ggsave("output/figures/autolysis_top3_boxplot.pdf", autolysis_plot, width = 12, height = 8, dpi = 300)
}

if(!is.null(pan_rin_plot)) {
  print(pan_rin_plot)
  ggsave("output/figures/pan_tissue_rin_scatter.pdf", pan_rin_plot, width = 10, height = 8, dpi = 300)
}

if(!is.null(pan_autolysis_plot)) {
  print(pan_autolysis_plot)
  ggsave("output/figures/pan_tissue_autolysis_boxplot.pdf", pan_autolysis_plot, width = 10, height = 8, dpi = 300)
}

#------------------------------------------------------------------------------
# Tissue-wise performance summary 
#------------------------------------------------------------------------------

create_tissue_performance_summary <- function(test_results) {
  # RIN performance by tissue
  rin_performance <- test_results[used_for_rin == TRUE & !is.na(rin_prediction), .(
    rin_r = cor(rin_prediction, true_rin),
    rin_rmse = sqrt(mean((rin_prediction - true_rin)^2)),
    rin_n = .N
  ), by = predicted_tissue]
  
  # Autolysis performance by tissue
  autolysis_performance <- test_results[used_for_autolysis == TRUE & !is.na(autolysis_prediction), .(
    autolysis_r = cor(autolysis_prediction, true_autolysis),
    autolysis_rmse = sqrt(mean((autolysis_prediction - true_autolysis)^2)),
    autolysis_accuracy = {
      pred_rounded <- pmax(0, pmin(3, round(autolysis_prediction)))
      mean(pred_rounded == true_autolysis)
    },
    autolysis_n = .N
  ), by = predicted_tissue]
  
  # Combine performance metrics
  tissue_summary <- merge(rin_performance, autolysis_performance, 
                         by = "predicted_tissue", all = TRUE)
  
  # Performance comparison plot
  performance_comparison <- ggplot(tissue_summary, aes(x = reorder(predicted_tissue, rin_r))) +
    geom_point(aes(y = rin_r, color = "RIN", size = rin_n), alpha = 0.7) +
    geom_point(aes(y = autolysis_r, color = "Autolysis", size = autolysis_n), alpha = 0.7) +
    coord_flip() +
    labs(x = "Tissue", y = "Correlation (R)", 
         title = "Tissue-Specific Performance Comparison",
         color = "Prediction Type", size = "Sample Size") +
    theme_minimal() +
    scale_color_manual(values = c("RIN" = "blue", "Autolysis" = "orange"))
  
  return(list(summary_table = tissue_summary, plot = performance_comparison))
}

# Generate remaining visualizations
tissue_performance <- create_tissue_performance_summary(test_results)

print(tissue_performance$plot)
ggsave("output/figures/tissue_performance_comparison.pdf", tissue_performance$plot, width = 12, height = 10, dpi = 300)

# Print summary table
cat("\nTissue Performance Summary:\n")
print(tissue_performance$summary_table[order(-rin_r)])
```




#------------------------------------------------------------------------------
# Pan-Tissue Figure for legends
#------------------------------------------------------------------------------
```{r}
make_spread_legend <- function(tissue_colors, n_rows = 2, point_size = 4, text_size = 4) {
  tissues <- names(tissue_colors)
  n <- length(tissues)
  
  # Wrap names at space for multi-word tissues
  tissue_labels <- str_replace_all(tissues, " ", "  \n")
  
  # Items per row
  per_row <- ceiling(n / n_rows)
  
  # Evenly spaced columns across fixed width
  total_width <- 100
  col_positions <- rep(seq(5, total_width - 5, length.out = per_row), times = n_rows)[1:n]
  row_labels <- rep(paste("Row", seq_len(n_rows)), each = per_row)[1:n]
  
  legend_data <- data.frame(
    tissue = tissues,
    label = tissue_labels,
    col = col_positions,
    row = factor(row_labels, levels = paste("Row", 1:n_rows))
  )
  
  ggplot(legend_data, aes(x = col, y = row)) +
    geom_point(aes(color = tissue), size = point_size) +
    geom_text(aes(label = label), size = text_size, hjust = -0.2, vjust = 0.5, color = "black", lineheight = 0.9) +
    scale_color_manual(values = tissue_colors, guide = "none") +
    scale_x_continuous(limits = c(0, total_width), expand = c(0, 0)) +
    scale_y_discrete(expand = expansion(mult = c(0.25, 0.25))) +  # more vertical space
    theme_void() +
    theme(
      plot.margin = margin(5, 10, 5, 10),
      plot.title = element_text(hjust = 0.5, size = 14, face = "bold")
    ) +
    coord_cartesian(clip = "off") +
    labs(title = "Tissue")
}

# Example usage
manual_legend <- make_spread_legend(tissue_colors, n_rows = 2, point_size = 4, text_size = 4)

plots_combined <- plot_grid(
  create_rin_plot(test_results, show_legend = FALSE),
  create_autolysis_plot(test_results, show_legend = FALSE),
   labels = c("B", "C"), label_size = 16,
  nrow = 1
)

final_with_manual_legend <- plot_grid(
  plots_combined,
  manual_legend,
  ncol = 1,
  rel_heights = c(1, 0.3)
)

print(final_with_manual_legend)

```


```{r}
#------------------------------------------------------------------------------
# Pan-Tissue Figure
#------------------------------------------------------------------------------

# Load pan-tissue model diagram
biorender_img <- magick::image_read("output/figures/pathqc Pan model.png", density = 600)[1]
biorender_plot <- ggdraw() + draw_image(biorender_img)

# Row 1: Panel A (Model diagram)
first_row <- plot_grid(
  biorender_plot,
  labels = "A",
  label_size = 16,
  rel_heights = c(0.9),
  ncol = 1
)

# Row 2: Panel B and C (RIN and Autolysis performance)
second_row <- plot_grid(
  final_with_manual_legend,
  rel_widths = c(1, 1),
  nrow = 1
)

# Combine all rows
combined_plot_final <- plot_grid(
  first_row,
  second_row,
  ncol = 1,
  rel_heights = c(1, 1.2),
  align = "v",
  axis = "lr",
  vjust = 0.5,
  hjust = 0.5
)

# Save final figure
ggsave("output/figures/PanTissue_Performance.pdf", combined_plot_final, width = 16, height = 10, dpi = 600)
```

#------------------------------------------------------------------------------
# Tissue Classification Visualization (Confusion Matrix + AUC)
#------------------------------------------------------------------------------t
```{r}
# 1. Tissue Classification Confusion Matrix
create_confusion_plot <- function(confusion_matrix) {
  cm_df <- as.data.frame(confusion_matrix)
  names(cm_df) <- c("Actual", "Predicted", "Count")
  
  ggplot(cm_df, aes(x = Actual, y = Predicted, fill = Count)) +
    geom_tile(color = "white") +
    geom_text(aes(label = Count), color = "black", size = 3) +
    scale_fill_gradient(low = "white", high = "steelblue") +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 10),
          axis.text.y = element_text(size = 10)) +
    labs(title = "Tissue Classification Confusion Matrix", 
         x = "True Tissue", y = "Predicted Tissue")
}
# Generate and save tissue classification plots
confusion_plot <- create_confusion_plot(performance$confusion_matrix)
print(confusion_plot)
ggsave("output/figures/tissue_confusion_matrix.pdf", confusion_plot, width = 14, height = 12, dpi = 300)
```

```{r}
# Pairwise AUC Matrix Implementation
calculate_pairwise_auc <- function(test_results) {
  tissues <- levels(test_results$true_tissue)
  n_tissues <- length(tissues)
  
  # Initialize matrix
  auc_matrix <- matrix(NA, nrow = n_tissues, ncol = n_tissues,
                       dimnames = list(tissues, tissues))
  
  # Calculate AUC for each tissue pair
  for(i in 1:n_tissues) {
    tissue_pos <- tissues[i]
    prob_col_pos <- paste0("prob_", gsub("[^A-Za-z0-9]", "_", tissue_pos))
    
    for(j in 1:n_tissues) {
      if(i == j) next  # Skip diagonal
      
      tissue_neg <- tissues[j]
      prob_col_neg <- paste0("prob_", gsub("[^A-Za-z0-9]", "_", tissue_neg))
      
      # Get data for this tissue pair
      pair_data <- test_results[true_tissue %in% c(tissue_pos, tissue_neg)]
      
      if(nrow(pair_data) > 0) {
        # Calculate relative probability
        prob_pos <- pair_data[[prob_col_pos]]
        prob_neg <- pair_data[[prob_col_neg]]
        relative_prob <- prob_pos / (prob_pos + prob_neg)
        
        # Binary labels
        binary_labels <- as.numeric(pair_data$true_tissue == tissue_pos)
        
        # Calculate AUC
        if(length(unique(binary_labels)) > 1) {
          roc_obj <- roc(binary_labels, relative_prob, quiet = TRUE)
          auc_matrix[i, j] <- as.numeric(auc(roc_obj))
        }
      }
    }
  }
  
  return(auc_matrix)
}

# Calculate matrix
auc_matrix <- calculate_pairwise_auc(test_results)

# Print matrix for inspection
cat("Pairwise AUC Matrix (29x29):\n")
print(round(auc_matrix, 3))

# Create heatmap
auc_melted <- melt(auc_matrix, na.rm = FALSE)
names(auc_melted) <- c("Target_Tissue", "Background_Tissue", "AUC")

# Summary of AUC range for better color scaling
auc_range <- range(auc_melted$AUC, na.rm = TRUE)
cat("AUC Range:", round(auc_range, 4), "\n")

pairwise_auc_plot <- ggplot(auc_melted, aes(x = Background_Tissue, y = Target_Tissue, fill = AUC)) +
  geom_tile(color = "white", size = 0.1) +
  # Optimized color scale for high AUC values (0.97-1.0 range)
  scale_fill_gradient(low = "lightblue", high = "darkblue", 
                      limits = c(0.99, 1.0), na.value = "grey90",
                      breaks = c(0.99, 0.995, 1.0),
                      labels = c("0.99", "0.995", "1.00")) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 10),
        axis.text.y = element_text(size = 10),
        legend.position = "right") +
  labs(title = "Pairwise AUC Matrix (Tissue Classification)", 
       x = "Background Tissue (Distinguished From)", 
       y = "Target Tissue (To Detect)",
       fill = "AUC") +
  coord_fixed()

print(pairwise_auc_plot)
ggsave("output/figures/pairwise_auc_heatmap.pdf", pairwise_auc_plot, width = 12, height = 10, dpi = 300)
```


```{r}
#------------------------------------------------------------------------------
# Figure S12: Pan-Tissue Model Performance Analysis
#------------------------------------------------------------------------------

# Left side: 2x2 grid (A, B, D, E)
left_grid <- plot_grid(
  rin_plot,                    # Panel A: Top 3 RIN scatter plots
  autolysis_plot,              # Panel B: Top 3 autolysis boxplots
  confusion_plot,              # Panel D: Confusion matrix  
  pairwise_auc_plot,          # Panel E: Pairwise AUC heatmap
  labels = c("A", "B", "D", "E"),
  label_size = 16,
  nrow = 2,
  ncol = 2,
  rel_widths = c(1, 1),
  rel_heights = c(1, 1)
)

# Right side: Panel C (full height)
right_panel <- plot_grid(
  tissue_performance$plot,     # Panel C: Cross-tissue performance comparison
  labels = "C",
  label_size = 16,
  ncol = 1
)

# Combine left grid with right panel
figure_s12_combined <- plot_grid(
  left_grid,
  right_panel,
  nrow = 1,
  ncol = 2,
  rel_widths = c(2, 1),        # Left side twice as wide as right
  align = "v"
)

# Save Figure S12
ggsave("output/figures/Figure_S12_PanTissue_Analysis.pdf", 
       figure_s12_combined, 
       width = 20, height = 12, dpi = 600, bg = "white")

print("Figure S12 created successfully!")
print(figure_s12_combined)
```
```{r}
#------------------------------------------------------------------------------
# Automatic Complete Performance Table Generation
#------------------------------------------------------------------------------
create_complete_performance_table <- function(rin_train_perf, auto_train_perf, test_results) {
  # Convert training performance to data.table if not already
  rin_train_perf <- as.data.table(rin_train_perf)
  auto_train_perf <- as.data.table(auto_train_perf)
  
  # Get test performance
  rin_test_perf <- test_results[used_for_rin == TRUE & !is.na(rin_prediction), .(
    test_rin_r = cor(rin_prediction, true_rin),
    test_rin_rmse = sqrt(mean((rin_prediction - true_rin)^2)),
    test_rin_n = .N
  ), by = predicted_tissue]
  
  autolysis_test_perf <- test_results[used_for_autolysis == TRUE & !is.na(autolysis_prediction), .(
    test_auto_r = cor(autolysis_prediction, true_autolysis),
    test_auto_rmse = sqrt(mean((autolysis_prediction - true_autolysis)^2)),
    test_auto_n = .N
  ), by = predicted_tissue]
  
  # Prepare training data (using standard data.frame operations)
  rin_train_clean <- data.table(
    tissue = rin_train_perf$tissue,
    train_rin_r = rin_train_perf$r_val,
    train_rin_rmse = rin_train_perf$rmse,
    train_rin_n = rin_train_perf$n
  )
  
  auto_train_clean <- data.table(
    tissue = auto_train_perf$tissue,
    train_auto_r = auto_train_perf$r_val,
    train_auto_rmse = auto_train_perf$rmse,
    train_auto_n = auto_train_perf$n
  )
  
  # Merge training performance
  train_combined <- merge(rin_train_clean, auto_train_clean, by = "tissue", all = TRUE)
  
  # Merge with test performance
  complete_table <- merge(train_combined, rin_test_perf, 
                         by.x = "tissue", by.y = "predicted_tissue", all.x = TRUE)
  complete_table <- merge(complete_table, autolysis_test_perf, 
                         by.x = "tissue", by.y = "predicted_tissue", all.x = TRUE)
  
  # Clean and format
  final_table <- complete_table[, .(
    Tissue = tissue,
    Train_RIN_R = round(train_rin_r, 3),
    Test_RIN_R = round(test_rin_r, 3),
    Train_Auto_R = round(train_auto_r, 3), 
    Test_Auto_R = round(test_auto_r, 3),
    Train_RIN_N = train_rin_n,
    Test_RIN_N = test_rin_n,
    Train_Auto_N = train_auto_n,
    Test_Auto_N = test_auto_n,
    Train_RIN_RMSE = round(train_rin_rmse, 3),
    Test_RIN_RMSE = round(test_rin_rmse, 3),
    Train_Auto_RMSE = round(train_auto_rmse, 3),
    Test_Auto_RMSE = round(test_auto_rmse, 3)
  )]
  
  # Order by training RIN performance (handle NA values)
  final_table <- final_table[order(-Train_RIN_R, na.last = TRUE)]
  
  return(final_table)
}

# Generate the complete table
complete_performance_table <- create_complete_performance_table(
  rin_training_performance, 
  autolysis_training_performance, 
  test_results
)

# Save as CSV
write.csv(complete_performance_table, 
          "output/figures/Supplementary_Table_S1_Complete_Performance.csv", 
          row.names = FALSE)

# Print preview
cat("Supplementary Table S1: Complete Training vs Test Performance\n")
cat("==============================================================\n")
print(complete_performance_table)

cat("\nTable saved to: output/figures/Supplementary_Table_S1_Complete_Performance.csv\n")
```



